## Step 1 - Collecting Data ---- 
The Happiness Scores and Ranking Data was collected from Kaggle.com from the following links:   
https://www.kaggle.com/unsdsn/world-happiness/downloads/2015.csv  
https://www.kaggle.com/unsdsn/world-happiness/downloads/2016.csv  
The scores are based on answers to the main life evaluation question asked in the poll. This question, known as the Cantril ladder, asks respondents to think of a ladder with the best possible life for them being a 10 and the worst possible life being a 0 and to rate their own current lives on that scale. The scores are from nationally representative samples for the years 2013-2015 and use the Gallup weights to make the estimates representative. The columns following the happiness score estimate the extent to which each of six factors ¨C economic production, social support, life expectancy, freedom, absence of corruption, and generosity ¨C contribute to making life evaluations higher in each country than they are in Dystopia, a hypothetical country that has values equal to the world¡¯s lowest national averages for each of the six factors. 

## Step 2: Exploring and preparing the data ---- 
## 2015 DATA ##
Import data and look into it, rename the attributes.
```{r}
# Sunny's working directory
#setwd("~/Babylam/CSUEB/2017 Spring/STAT 6509/project")
# Sherry's working directory
#setwd("~/Documents/6509")
#Zhaoyan's working directory
setwd("C:/Users/document")
raw <- read.csv("2015.csv")
summary(raw)
library(readxl)
happy15_original = read_excel("2015.xlsx")
#str(happy15_original)
happy15_ml = happy15_original[,c(1,5:13)]
colnames(happy15_ml) <- c("Region","HScore", "LogGDP", "socialSupport", "Health", "Freedom", "Generosity","GovernmentTrust","positiveAffect","negativeAffect")
summary(happy15_ml[,-c(1)])
happy15_ml = na.exclude(happy15_ml)
happy15 = happy15_ml[,-c(1)]
#str(happy15)
summary(happy15)
```



Diagnostics for predictor variables
```{r}
par(mfrow=c(2,2))
boxplot(happy15$LogGDP, xlab = "LogGDP")
hist(happy15$LogGDP, xlab = "LogGDP",main ="")
boxplot(happy15$socialSupport, xlab = "socialSupport")
hist(happy15$socialSupport, xlab = "socialSupport",main ="")
boxplot(happy15$Health, xlab = "Health")
hist(happy15$Health, xlab = "Health",main ="")
boxplot(happy15$Freedom, xlab = "Freedom")
hist(happy15$Freedom, xlab = "Freedom",main ="")
boxplot(happy15$GovernmentTrust, xlab = "GovernmentTrust")
hist(happy15$GovernmentTrust, xlab = "GovernmentTrust",main ="")
boxplot(happy15$Generosity, xlab = "Generosity")
hist(happy15$Generosity, xlab = "Generosity",main ="")
boxplot(happy15$positiveAffect, xlab = "positiveAffect")
hist(happy15$positiveAffect, xlab = "positiveAffect",main ="")
boxplot(happy15$negativeAffect,xlab="negativeAffect")
hist(happy15$negativeAffect,xlab = "negativeAffect",main ="")
```



Correlation test between Happiness score and all other variables. Visualization of correlation between variables and Happiness score. 
```{r}
# eliminate Country name, region, happiness rank, lower & upper confidence interval
library(psych)
print(corr.test(happy15, y = NULL, use = "pairwise",method="pearson",adjust="holm",alpha=.05), short = FALSE)
library(corrplot)
# Correlation plot to visualize data
cormatrix <- cor(as.matrix(happy15))
corrplot(cormatrix,method = "square")
library(psych)
pairs.panels(data.frame(happy15$HScore,happy15$LogGDP,happy15$socialSupport,happy15$Health,happy15$Freedom,happy15$GovernmentTrust,happy15$Generosity,happy15$positiveAffect,happy15$negativeAffect))
```
AIC test
```{r}
# from 1 to all interactions
null <- lm(happy15$HScore~1,data = happy15)
full <- lm(happy15$HScore~.^2,data = happy15)
step(null,scope = list(lower = null,upper =full),direction = 'forward')
```

Compare models:  
```{r}
# lowest AIC model
# model AIC: HScore =LogGDP + positiveAffect + socialSupport + Freedom + Health + LogGDP*positiveAffect + positiveAffect*Health + socialSupport*Freedom
happy15_lmAIC = lm(HScore ~ LogGDP + positiveAffect + socialSupport + Freedom + 
    Health + LogGDP*positiveAffect + positiveAffect*Health + socialSupport*Freedom,data = happy15)
summary(happy15_lmAIC)
# additive model with ALL interactions
# model 1: HScore = LogGDP+socialSupport+Health+Freedom+GovernmentTrust+Generosity+positiveAffect+negativeAffect+LogGDP*socialSupport+LogGDP*Health+LogGDP*Freedom+LogGDP*GovernmentTrust+LogGDP*Generosity+LogGDP*positiveAffect+LogGDP*negativeAffect+socialSupport*Health+socialSupport*Freedom+socialSupport*GovernmentTrust+socialSupport*Generosity+socialSupport*positiveAffect+socialSupport*negativeAffect+Health*Freedom+Health*GovernmentTrust+Health*Generosity+Health*positiveAffect+Health*negativeAffect+GovernmentTrust*Generosity+GovernmentTrust*positiveAffect+GovernmentTrust*negativeAffect+positiveAffect*negativeAffect [8 factors with interactions]
happy15_lm1 = lm(HScore~.^2, data=happy15)
summary(happy15_lm1)
# additive model with ALL significant correlation predictors
# model 2: HScore = LogGDP+socialSupport+Health+Freedom+GovernmentTrust+Generosity+positiveAffect+negativeAffect+LogGDP*socialSupport+LogGDP*Health+LogGDP*Freedom+LogGDP*GovernmentTrust+LogGDP*positiveAffect++LogGDP*negativeAffect+socialSupport*Health+socialSupport*Freedom+socialSupport*GovernmentTrust+socialSupport*Generosity+socialSupport*positiveAffect+socialSupport*negativeAffect+Health*Freedom+Health*GovernmentTrust+Health*positiveAffect+Health*negativeAffect+Freedom*GovernmentTrust+Freedom*Generosity+Freedom*positiveAffect+Freedom*negativeAffect+GovernmentTrust*Generosity+GovernmentTrust*positiveAffect+GovernmentTrust*negativeAffect+Generosity*positiveAffect+positiveAffect*negativeAffect
happy15_lm2 = lm(HScore~LogGDP+socialSupport+Health+Freedom+GovernmentTrust+Generosity+positiveAffect+negativeAffect+
                LogGDP*socialSupport+LogGDP*Health+LogGDP*Freedom+LogGDP*GovernmentTrust+LogGDP*positiveAffect+LogGDP*negativeAffect+
                   socialSupport*Health+socialSupport*Freedom+socialSupport*GovernmentTrust+socialSupport*positiveAffect+socialSupport*negativeAffect+
                   Health*Freedom+Health*GovernmentTrust+Health*positiveAffect+Health*negativeAffect+Freedom*GovernmentTrust+Freedom*Generosity+Freedom*positiveAffect+Freedom*negativeAffect+
                   Generosity*GovernmentTrust+Generosity*positiveAffect+Generosity*negativeAffect+
                   GovernmentTrust*positiveAffect+GovernmentTrust*negativeAffect+positiveAffect*negativeAffect,data = happy15)
summary(happy15_lm2)
anova(happy15_lm1,happy15_lm2)
# additive model
# model 3: HScore = LogGDP+socialSupport+Health+Freedom+GovernmentTrust+Generosity+positiveAffect+negativeAffect [8 factors]
happy15_lm3 = lm(HScore~LogGDP+socialSupport+Health+Freedom+GovernmentTrust+Generosity+positiveAffect+negativeAffect,data = happy15)
summary(happy15_lm3)
anova(happy15_lm2,happy15_lm3)
# additive model - freedom - generosity - negative effect
# model 4: HScore = LogGDP+socialSupport+Health+GovernmentTrust+positiveAffect  [5 factors]
happy15_lm4 = lm(HScore~LogGDP+socialSupport+Health+GovernmentTrust+positiveAffect,data = happy15)
summary(happy15_lm4)
anova(happy15_lm3,happy15_lm4)
# additive model - freedom - generosity
# model 5: HScore = LogGDP+socialSupport+Health+GovernmentTrust+positiveAffect+negativeAffect  [6 factors]
happy15_lm5 = lm(HScore~LogGDP+socialSupport+Health+GovernmentTrust+positiveAffect+negativeAffect,data = happy15)
summary(happy15_lm5)
anova(happy15_lm3,happy15_lm5)
# final model is happy15_lm5
happy15_lm = happy15_lm5
```

From comparing model, the model we are going to used is  
HScore = LogGDP+socialSupport+Health+Freedom+GovernmentTrust+Generosity+DystopiaRes  
Residual diagnosis
```{r}
# plot residuals
par(mfrow=c(2,2))
plot(happy15_lm)
par(mfrow=c(1,1))
# boxplot of residuals
#boxplot(happy15_lm$residuals, xlab = "residuals")
# histogram for residuals 
#hist(happy15_lm$residuals, xlab = "residuals")
# normality test for residual => REJECT normality of error variance
shapiro.test(happy15_lm$residuals)
# residual plot
library(car)
residualPlots(happy15_lm)
# Breush-Pagan test => REJECT constancy of error variance
library(lmtest)
bptest(happy15_lm)
ncvTest(happy15_lm)
# marginal model plots
marginalModelPlots(happy15_lm)
# qqPlot
qqPlot(happy15_lm, id.n = 2)
# outlierTest
outlierTest(happy15_lm)
# infludenceIndexPlot
influenceIndexPlot(happy15_lm, id.n=2)
# influencePlot
influencePlot(happy15_lm)
```

World map to visualized the data
```{r}
library(ggplot2)
library(dplyr)
library(highcharter)
library(formattable)
library(plotly)
library(countrycode)
library(maps)
library(readr)
X2015 <- read_csv("2015.csv")
summary(X2015)
happy15_map <- X2015[,c(1,4)]
colnames(happy15_map) <- c("Country", "HScore")
# correct country name from database
happy15_map$Country<-as.character(happy15_map$Country)
correction15<-c("Congo (Brazzaville)"="Democratic Republic of the Congo","Guinea"="Equatorial Guinea","North Cyprus"="Northern Cyprus","Serbia"="Republic of Serbia","Congo (Brazzaville)"="Republic of Congo","Somaliland region"="Somaliland","Tanzania"="United Republic of Tanzania","United States Minor Outlying Islands"="United States","United States"="United States of America","United States"="United States Virgin Islands","Russia"="Russian Federation","Venezuela"="Venezuela, Bolivarian Republic of","Bolivia"="Bolivia (Plurinational State of)" )
for(i in names(correction15)){
  happy15_map[happy15_map$Country==i,"Country"]<-correction15[i]
}
# match country name along iso3 code
countrycode_data$country.name.en<-as.factor(countrycode_data$country.name.en)
happy15_map$iso3<-countrycode_data[match(happy15_map$Country,countrycode_data$country.name.en),"iso3c"]
# create the world map
map15 <- highchart() %>% 
  hc_add_series_map(worldgeojson, happy15_map, value = "HScore", joinBy = "iso3") %>% 
  hc_colorAxis(min=min(happy15_map$HScore),max=max(happy15_map$HScore),minColor = "#FACFCF", maxColor = "#7569FD")%>%
  hc_mapNavigation(enabled = TRUE) %>%
  hc_credits(enabled = F) 
  hc_title(text = "2015 Happiness Report World Map", style = list(color = "#2b908f", fontWeight = "bold"), map15)
## 2016 DATA ##
X2016 <- read_csv("2016.csv")
happy16_map <- X2016[,c(1,4)]
colnames(happy16_map) <- c("Country", "HScore")
# correct country name from database
happy16_map$Country<-as.character(happy16_map$Country)
correction16<-c("Congo (Brazzaville)"="Democratic Republic of the Congo","Guinea"="Equatorial Guinea","North Cyprus"="Northern Cyprus","Serbia"="Republic of Serbia","Congo (Brazzaville)"="Republic of Congo","Somaliland region"="Somaliland","Tanzania"="United Republic of Tanzania","United States Minor Outlying Islands"="United States","United States"="United States of America","United States"="United States Virgin Islands","Somaliland Region"="Somaliland","Russia"="Russian Federation","Venezuela"="Venezuela, Bolivarian Republic of","Bolivia"="Bolivia (Plurinational State of)" )
for(i in names(correction16)){
  happy16_map[happy16_map$Country==i,"Country"]<-correction16[i]
}
# match country name along iso3 code
countrycode_data$country.name.en<-as.factor(countrycode_data$country.name.en)
happy16_map$iso3<-countrycode_data[match(happy16_map$Country,countrycode_data$country.name.en),"iso3c"]
# create the world map
map16 <- highchart() %>% 
  hc_add_series_map(worldgeojson, happy16_map, value = "HScore", joinBy = "iso3") %>% 
  hc_colorAxis(min=min(happy16_map$HScore),max=max(happy16_map$HScore),minColor = "#FACFCF", maxColor = "#7569FD")%>%
  hc_mapNavigation(enabled = TRUE) %>%
  hc_credits(enabled = F) 
  hc_title(text = "2016 Happiness Report World Map", style = list(color = "#2b908f", fontWeight = "bold"), map16)
```

Box-Plots of happiness score in different regions
```{r}
## 2015 data ##
plot_ly(X2015,x = X2015$Region,y=X2015$`Happiness Score`,
        type = "box", boxpoints ="all",pointpos = -1.8, color = X2015$Region)%>%
  layout(xaxis=list(showticklabels = FALSE),margin = list(b=100),title ="2015 Happiness.Score by Region")    
## 2016 data ##
plot_ly(X2016,x = X2016$Region,y=X2016$`Happiness Score`,
        type = "box", boxpoints ="all",pointpos = -1.8, color = X2016$Region)%>%
  layout(xaxis=list(showticklabels = FALSE),margin = list(b=100),title ="2016 Happiness.Score by Region")
```


input data from 2016
```{r}
happy16_original = read_excel("2016.xlsx")
str(happy16_original)
happy16_ml = happy16_original[,c(1,4:12)]
colnames(happy16_ml) <- c("Region","HScore", "LogGDP", "socialSupport", "Health", "Freedom", "Generosity","GovernmentTrust","positiveAffect","negativeAffect")
happy16_ml = na.exclude(happy16_ml)
str(happy16_ml)
```



Correlation test between Happiness score and all other variables. Visualization of correlation between variables and Happiness score. Result is similar to 2016 data.
```{r}
# eliminate Country name, region, happiness rank, lower & upper confidence interval
print(corr.test(happy16_ml[,-c(1)], y = NULL, use = "pairwise",method="pearson",adjust="holm",alpha=.05), short = FALSE)
# Correlation plot to visualize data
cormatrix <- cor(as.matrix(happy16_ml[,-c(1)]))
corrplot(cormatrix,method = "square")
pairs.panels(data.frame(happy16_ml$HScore,happy16_ml$LogGDP,happy16_ml$socialSupport,happy16_ml$Health,happy16_ml$Freedom,happy16_ml$GovernmentTrust,happy16_ml$Generosity,happy16_ml$positiveAffect,happy16_ml$negativeAffect))
```

Compare data between 2015 and 2016 data. 
```{r}
# based on regression model above, only use the following factors to do classification
# HScore, LogGDP, socialSupport, Health, GovernmentTrust, positiveAffect, negativeAffect
happy15_ml <- happy15_ml[,-c(6,7)]
happy16_ml <- happy16_ml[,-c(6,7)]
happy15_ml$Region <- factor(happy15_ml$Region, levels = c("Australia and New Zealand","Central and Eastern Europe","Eastern Asia","Latin America and Caribbean","Middle East and Northern Africa","North America","Southeastern Asia","Southern Asia","Sub-Saharan Africa","Western Europe"), labels = c("Australia and New Zealand","Central and Eastern Europe","Eastern Asia","Latin America and Caribbean","Middle East and Northern Africa","North America","Southeastern Asia","Southern Asia","Sub-Saharan Africa","Western Europe"))
str(happy15_ml)
happy16_ml$Region <- factor(happy16_ml$Region, levels = c("Australia and New Zealand","Central and Eastern Europe","Eastern Asia","Latin America and Caribbean","Middle East and Northern Africa","North America","Southeastern Asia","Southern Asia","Sub-Saharan Africa","Western Europe"), labels = c("Australia and New Zealand","Central and Eastern Europe","Eastern Asia","Latin America and Caribbean","Middle East and Northern Africa","North America","Southeastern Asia","Southern Asia","Sub-Saharan Africa","Western Europe"))
str(happy16_ml)
round(prop.table(table(happy15_ml$Region))*100, digits = 1)
round(prop.table(table(happy16_ml$Region))*100, digits = 1)
```


```{r}
# Step 3: Use C5.0 model to predict 2016 data ----
# Normalize data
normalize <- function(x) {
  return((x-min(x))/ (max(x)-min(x)))
}
happy15_ml_n <- as.data.frame(lapply(happy15_ml[2:8],normalize))
happy16_ml_n <- as.data.frame(lapply(happy16_ml[2:8],normalize))
summary(happy15_ml_n)
summary(happy16_ml_n)
# create training and test dataset
happy_train <- happy15_ml_n
happy_test <- happy16_ml_n
happy_train_label <- happy15_ml$Region
happy_test_label <- happy16_ml$Region
# training a C5.0 model on the data
library(C50)
happy_model <- C5.0(happy_train, happy_train_label)
summary(happy_model)
# Step 4: Evaulate the prediction throught residuals performance
happy_pred <- predict(happy_model, happy_test)
# cross tabulation of predicted versus actual classes
# accuracy rate = (20+3+12+3+30+19)/121= 87/121 = 71.90 %
library(gmodels)
CrossTable(happy_test_label, happy_pred, prop.chisq = FALSE, prop.c = FALSE, prop.r = FALSE, dnn = c('actual default', 'predicted default'))
# Step 5: boosted the accuracy of decision tree with 10 trials
happy_model_boost10 <- C5.0(happy_train, happy_train_label, trials = 20)
summary(happy_model_boost10)
happy_pred_boost10 <- predict(happy_model_boost10, happy_test)
# accuracy rate = (20+3+12+3+3+30+19)/121= 90/121 = 74.38 %
CrossTable(happy_test_label, happy_pred, prop.chisq = FALSE, prop.c = FALSE, prop.r = FALSE,dnn = c('actual default', 'predicted default'))
```


## linear regression prediction ##
```{r}
happy_pred_linear <- predict(happy15_lm,happy16_ml)
happy_pred_residual = happy_pred_linear-happy16_ml$HScore
summary(happy_pred_linear)
summary(happy16_ml$HScore)
summary(happy_pred_residual)
plot(happy_pred_linear, happy16_ml$HScore)
abline(0,1)
plot(happy_pred_residual)
abline(0,0)
# Mean Absolute Error
MAE <- function(actual,predicted) {mean(abs(actual-predicted))}
MAE(happy16_ml$HScore, happy_pred_linear)
```
```


